---
title: "Report_motif_analysis"
author: "Charlotte Vandermeulen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("rmarkdown")
library("tidyverse")
#Rscript -e "rmarkdown::render(script_path,params=list(directory = path_of_pipeline))"
#knitr::opts_knit$set(root.dir = "/mnt/sharc/mnt/fastdata/bo1cv/a549_slam/strand_one")
```

# Report of pipeline motif_analysis

In this report, we summarize the motifs discovered by the pipeline.

"High" stability motifs refer to motifs that were enriched in the 3'UTR of 
transcripts ranking at the top for the value used to rank them. Since originally
the pipeline was used on mRNA half-lives, I gave them the name of "highstab" motifs.
Similarly, "low" stability motifs refer to motifs enriched in the bottom 
transcripts for the value used to rank them.
Other ranking values can be used (such as Translation Efficiency, ...), keeping 
in mind that the pipeline will label higher values as high "stability", "TE", ...

It reports: <br>
- the number of motifs discovered by Homer, Streme and/or Fire; <br>
- the number of motifs that contained/matched miRNA seed targets and 
consensus polyA sites (AUAAA or AAUAAA); <br>
- for highstab motifs, the number of motifs that matched lowstab 
motifs (the final list of highstab motifs has been filtered for 
these, while lowstab motifs have been filetred only for polyAs); <br>
- which motif enrichment tool discovered the different motifs, Streme, Homer 
and/or Fire; <br>
- which "conditions" (ex: half-life, TE, ...).
 
N.B. Some motifs may have been discovered by different tools/conditions, if 
it is the case, the multiple source column will have several entries (cf tables
below).



```{r loading, include=FALSE}
##Streme and Homer
#Get list of motifs from Streme 
getwd()
files_streme = list.files(pattern = "streme.txt",recursive = T) 
files_streme <- files_streme[!(grepl("tomtom.log", files_streme))]
streme <- sapply(files_streme, readLines)
streme <- lapply(streme, function(x) x[grep("MOTIF", x)])
streme <- lapply(streme, data.frame)
#lapply(streme, function(x) separate_rows(x[1], sep = " "))
streme <- bind_rows(streme, .id = "source")
streme$X..i.. <- str_remove_all(streme$X..i.., pattern = "MOTIF | STREME-\\d+| ")
colnames(streme) <- c("source","motif_name")
streme$source <- str_remove(streme$source, ".dir/streme.txt")

#Get list of motifs from Streme 
files_homer = list.files(pattern = "homerMotifs.all.motifs.meme",recursive = T) 
files_homer <- files_homer[!(grepl("tomtom.log", files_homer))]
homer <- sapply(files_homer, readLines)
homer <- lapply(homer, function(x) x[grep("MOTIF", x)])
homer <- lapply(homer, data.frame)
homer <- bind_rows(homer, .id = "source")
homer$X..i.. <- str_remove_all(homer$X..i.., pattern = "MOTIF | HOMER-\\d+| ")
colnames(homer) <- c("source","motif_name")
homer$source <- str_remove(homer$source, ".dir/homerMotifs.all.motifs.meme")

#Get list of motifs merge
files_merge = list.files(pattern = "_merge_homer_streme.meme",recursive = T) 
files_merge <- files_merge[!(grepl(".log", files_merge))]
merge_homer_streme <- sapply(files_merge, readLines)
merge_homer_streme <- lapply(merge_homer_streme, function(x) x[grep("MOTIF", x)])
merge_homer_streme <- lapply(merge_homer_streme, data.frame)
merge_homer_streme <- bind_rows(merge_homer_streme, .id = NULL)
merge_homer_streme$X..i.. <- str_remove_all(merge_homer_streme$X..i.., pattern = "MOTIF | HOMER-\\d+| STREME-\\d+| ")
merge_homer_streme <- merge_homer_streme$X..i..

#Keep track of the ones that merged together ones that merged together
highstab_tomtom_streme_homer <- try(read.table("final_motifs/highstab_merge_homer_streme.motifs.tomtom", fill = T, header = T))
  #Check if tomtom file is empty
if (class(highstab_tomtom_streme_homer) != "try-error" | (any(is.na(highstab_tomtom_streme_homer)) == F) ) {
  highstab_tomtom_streme_homer <- highstab_tomtom_streme_homer[1:(nrow(highstab_tomtom_streme_homer)-4),1:2]

  #Loop merged motif in tomtom 
  #List of merged motifs
list_merged_high <- list()
  for (x in merge_homer_streme) {
    if (x %in% c(highstab_tomtom_streme_homer[,1], highstab_tomtom_streme_homer[,2])) {
      together <- highstab_tomtom_streme_homer %>%
        filter_all(any_vars(. %in% x))
      together <- c(together[,1], together[,2])
      list_merged_high <- append(list_merged_high, list(together))
    }
  }
} else {list_merged_high <- list()}
  
lowstab_tomtom_streme_homer <- try(read.table("final_motifs/lowstab_merge_homer_streme.motifs.tomtom", fill = T, header = T))

#Check if tomtom file is empty
if (class(lowstab_tomtom_streme_homer) != "try-error" & (any(is.na(lowstab_tomtom_streme_homer)) == F) ) {
  lowstab_tomtom_streme_homer <- lowstab_tomtom_streme_homer[1:(nrow(lowstab_tomtom_streme_homer)-4),c(1.2)]

  #Loop merged motif in tomtom 
  #List of merged motifs
  list_merged_low <- list()
  for (x in merge_homer_streme) {
    if (x %in% c(highstab_tomtom_streme_homer[,1], highstab_tomtom_streme_homer[,2])) {
      together <- highstab_tomtom_streme_homer %>%
        filter_all(any_vars(. %in% x))
      together <- c(together[,1], together[,2])
      list_merged_low <- append(list_merged_low, list(together))
    }
  }
} else {list_merged_low <- list()}


##Fire
files_fire = list.files(pattern = "allkmer.signif.motifs",recursive = T) 
fire_source <- sapply(files_fire, readLines)
fire_source <- lapply(fire_source, data.frame)
fire_source <- bind_rows(fire_source, .id = "source")
if (nrow(fire_source) > 0 ) {
  fire_source[c("motif_name", "sequence")]  <- str_split_fixed(fire_source$X..i.., "\\t",2)
  fire_source$X..i.. <- NULL
  fire_source$source <- str_remove_all(fire_source$source, pattern = "fire.dir/|.allkmer.signif.motifs| ")
  fire_source$source <- paste0(fire_source$source, "_fire")
  #Getting list of merges motifs high and low 
  fire_high <- fire_source %>% filter(str_detect(source, "high"))
  fire_low <- fire_source %>% filter(str_detect(source, "low"))
  
  #If no motifs in high
if (nrow(fire_high) == 0){
  print("No high motifs from fire")
} else {
  self_match <- sapply(fire_high$sequence, function(i) fire_high[str_detect(fire_high$sequence, i), 2])
self_match <- self_match[lapply(self_match,length) >1]
#Merge list of merged Homer/streme and Fire
list_merged_high <- c(list_merged_high, self_match)
}


if (nrow(fire_low) == 0){
  print("No low motifs from fire")
} else {
  self_match <- sapply(fire_low$sequence, function(i) fire_low[str_detect(fire_low$sequence, i), 2])
  self_match <- self_match[lapply(self_match,length) >1]
  #Merge list of merged Homer/streme and Fire
  list_merged_low <- c(list_merged_low, self_match)
}

#Finally, merge Homer, streme and Fire
all_possible_motifs_high <- bind_rows(homer, streme, .id = NULL) %>%
  bind_rows(., fire_source[,1:2], .id = NULL) %>%
  filter(grepl("high", source))
all_possible_motifs_low <- bind_rows(homer, streme, .id = NULL) %>%
  bind_rows(., fire_source[,1:2], .id = NULL) %>%
  filter(grepl("low", source))

} else {
  
  #Finally, merge Homer and homer (fire empty)
all_possible_motifs_high <- bind_rows(homer, streme, .id = NULL) %>%
  filter(grepl("high", source))
all_possible_motifs_low <- bind_rows(homer, streme, .id = NULL) %>%
  filter(grepl("low", source))
  
}



```


```{r create multiple table, include=FALSE}
#Loading final table of the determining multiple source motifs
highstab_final <- read.table("final_motifs/highstab_final_motifs.list",
                              header = T)

highstab_final_source <- c()
for (m in highstab_final$name) {
  its_name <- all_possible_motifs_high[all_possible_motifs_high$motif_name == m , 1]
  its_merge <- Filter(function(x) m %in% x, list_merged_high) %>%
    unlist() %>% unique()
  #print(paste0("any other: ", its_merge))
  
  its_merge_names <- lapply(its_merge, 
                            function(x)  
                              all_possible_motifs_high[all_possible_motifs_high$motif_name == x , 1 ]) %>%
    unlist() %>%  c(., its_name)  %>% unique()
  highstab_final_source <- c(highstab_final_source, paste(its_merge_names, collapse = ", "))
}
highstab_final$multiple_source <- highstab_final_source
#Some got a .1 I don't know why?? need to remove possible ones
highstab_final$name <- str_remove(highstab_final$name, "\\.\\d+")


lowstab_final <- read.table("final_motifs/lowstab_final_motifs.list", 
                            header = T)
lowstab_final$name <- str_remove(lowstab_final$name, "\\.\\d+")


lowstab_final_source <- c()
for (m in lowstab_final$name) {
  print("round")
  its_name <- all_possible_motifs_low[all_possible_motifs_low$motif_name == m , 1]
  its_merge <- Filter(function(x) m %in% x, list_merged_low) %>%
    unlist() %>% unique()
  #print(paste0("any other: ", its_merge))
  
  its_merge_names <- lapply(its_merge, 
                            function(x)  
                              all_possible_motifs_low[all_possible_motifs_low$motif_name == x , 1 ]) %>%
    unlist() %>%  c(., its_name)  %>% unique()
  lowstab_final_source <- c(lowstab_final_source, paste(its_merge_names, collapse = ", "))
}

lowstab_final$multiple_source <- lowstab_final_source
```

## "Highstability" motifs

The pipeline discovered `r nrow(highstab_final)` "high" stability motifs.

```{r highstablog, echo=FALSE}
readLines("final_motifs/highstab_final_motifs.list.log")
```
(from file final_motifs/highstab_final_motifs.list.log)


Table of high stability motifs and their origins: 
```{r highstab, echo=FALSE}
highstab_final
```

## "Lowstability" motifs

The pipeline discovered `r nrow(lowstab_final)` "low" stability motifs.


```{rlowstablog, echo=FALSE}
readLines("final_motifs/lowstab_final_motifs.list.log")
```
(from file final_motifs/lowstab_final_motifs.list.log)


Table of low stability motifs and their origins: 
```{r lowstab, echo=FALSE}
lowstab_final
```


Both these tables have been created in final_motifs as 
final_motifs/low/highstab_final_motifs.list.source.

If you would like more details on the different outputs genrated by the pipeline,
see the readme file.


```{r output, include= F}
write.table(highstab_final,
            "final_motifs/highstab_final_motifs.list.source", 
            col.names = T,
            row.names = F)

write.table(lowstab_final, 
            "final_motifs/lowstab_final_motifs.list.source", 
            col.names = T,
            row.names = F)
```


